---
title: "CADUCEUS Preliminary Results"
subtitle: "Presentation at Kohler, April 2018"
author: "Jennifer Thompson, MPH; Supervisor: Rameela Chandrasekhar, PhD"
output:
  html_notebook:
    theme: flatly
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
---

CADUCEUS is an ancillary study enrolling patients from the MENDS2, MOSAIC, and
INSIGHT studies in progress at Vanderbilt University Medical Center, looking at
three enzyme levels (BChE, AChE, and AChe in hemoglobin) during the course of
critical illness. This preliminary analysis describes study results as of April
2018.

```{r setup, results = "hide"}
knitr::opts_chunk$set(message = FALSE)

suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(glue))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(captioner))
suppressPackageStartupMessages(library(skimr))
suppressPackageStartupMessages(library(patchwork))
  ## Github only, till ggplot2 goes to CRAN :(
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(ggridges))

## Hmisc setup for nicely printing summaryM output
mu <- markupSpecs$html

## -- General helper functions (summary stats, formatting/printing) ------------
factor_tf <-
  function(x){ factor(as.numeric(x), levels = 1:0, labels = c("Yes", "No")) }
format_comma <- partial(format, big.mark = ",")
sum_na <- partial(sum, na.rm = TRUE)
q25 <- partial(quantile, probs = 0.25, na.rm = TRUE)
q50 <- partial(quantile, probs = 0.50, na.rm = TRUE)
q75 <- partial(quantile, probs = 0.75, na.rm = TRUE)
mean_na <- partial(mean, na.rm = TRUE)
sd_na <- partial(sd, na.rm = TRUE)
rndformat <- function(x, digits = 2){ format(round(x, digits), nsmall = digits) }
get_npct <- function(num, denom){
  sprintf("%s (%s%%)", num, round((num / denom) * 100))
}
describe_cont <- function(
  v_q50, v_q25, v_q75, v_mean, v_sd, dig_iqr = 0, dig_msd = 2
){
  sprintf(
    "%s [%s, %s]\\\n%s +/- %s",
    rndformat(v_q50, digits = dig_iqr),
    rndformat(v_q25, digits = dig_iqr),
    rndformat(v_q75, digits = dig_iqr),
    rndformat(v_mean, digits = dig_msd),
    rndformat(v_sd, digits = dig_msd)
  )
}

## Wrapper with my preferred options for printing summaryM objects
my_html <- function(sMobj, caption){
  html(
    sMobj,
    exclude1 = FALSE, long = TRUE, digits = 2, what = "%", npct = "both",
    prmsd = TRUE, brmsd = TRUE, middle.bold = TRUE,
    ## These options don't seem to be working.
    msdsize = mu$smaller2, outer.size = mu$smaller2, rowsep = TRUE,
    caption = caption
  )
}

## kable_styling wrapper to ensure all tables are consistently styled
mykablestyle <- function(obj, stripes = FALSE, ...){
  boptions <- c("hover", "responsive", "condensed", "bordered")
  if(stripes){ boptions <- c(boptions, "striped") }
  
  kable_styling(
    obj,
    bootstrap_options = boptions,
    full_width = FALSE,
    ...
  )
}

```

```{r captions, results = "hide"}
## -- Table and figure captions ------------------------------------------------
table_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")

table_nums(name = "descriptives", caption = "Description of CADUCEUS Cohort")

# fig_nums(name = "exclusions", caption = "Study Exclusions")

```

```{r load_data}
## -- Source script that reads in, combines current data from all studies ------
source("combine_data.R")

## -- Restrict cohort to patients, days with >=1 enzyme measurement ------------
daily_df <- filter(
  daily_df,
  !is.na(cad_ache_ul) | !is.na(cad_ache_ughb) | !is.na(cad_bche_ul)
)

# ## What study days have enzymes?
# ggplot(data = daily_df, aes(x = cad_day)) +
#   geom_bar(stat = "count") +
#   labs(
#     x = "Study Day (CADUCEUS Date - Enrollment Date)",
#     y = "Number of Enzyme Measurements"
#   )

## We want to focus on study days 1/3/5/7/14. These study days should have blood
## draws often, per the various study protocols, and thus have the most enzyme
## measurements. Other blood draws occur at the time of hospital discharge,
## which varies per patient and thus is less helpful for our current purposes.

## Due to common data collection happenings, combine day 2 with day 1; 4 with 3;
##  6 with 5
daily_sub <- daily_df %>%
  mutate(
    cad_day = ifelse(cad_day <= 6 & cad_day %% 2 == 0, cad_day - 1, cad_day),
    ## Dummy variable for plotting
    dummy = 1
  ) %>%
  filter(cad_day %in% c(seq(1, 7, 2), 14))

## Vector of unique patients with >=1 enzyme measurement
cad_pts <- unique(daily_sub$id)

## -- Add labels to descriptive dataset ----------------------------------------
oneobs_desc <- subset(oneobs_df, id %in% cad_pts)
label(oneobs_desc$age) <- "Age at enrollment"
label(oneobs_desc$gender) <- "Gender"
label(oneobs_desc$days_del) <- "Days with delirium"
label(oneobs_desc$days_coma) <- "Days with coma"
label(oneobs_desc$died_inhosp) <- "In-hospital mortality"

## Dummy variable, because formatting
oneobs_desc$allpts <- "All Patients"

```

# Description of Cohort

## Definition and Enzyme Time Points

We currently have `r length(cad_pts)` patients, enrolled in three studies, with
CADUCEUS enzyme measurements available on at least one of study days 1, 3, 5, 7,
and/or 14.[^1] We focus on these days because the three study protocols require
blood draws at subsets of those time points, and thus most of our enzyme levels
are clustered on those days.[^2] (Levels on study days 2, 4, and 6 were
clustered with the previous day's time point, in recognition of study protocols
and common data collection experiences.)

[^1]: "Study day" = days after enrollment; study day 1 = day of enrollment.
[^2]: Enzyme levels are available at other time points, generally because some
patients also contributed blood levels at hospital discharge, which obviously
varies per individual. Due to these varying timelines and low sample size on any
given day, these are less helpful for our current purposes and are not included.

## Baseline and In-Hospital Characteristics

`r table_nums("descriptives", display = "cite")` provides a basic description of
these patients at study enrollment as well as information on their mental status
throughout the study. **Note** that because our mortality rates among this
critically ill cohort are high, durations of delirium and coma are often
truncated by death.

```{r descriptives}
html(
  summaryM(
    age + gender + days_del + days_coma + died_inhosp ~ allpts,
    data = oneobs_desc
  ),
  long = TRUE,
  exclude1 = FALSE,
  what = "%",
  digits = 2,
  npct = "both",
  caption = "Cohort Baseline and In-Hospital Characteristics",
  insert.bottom = "Note that because data collection is ongoing, hospital mortality status is not yet available on all patients."
)

```

# Description of Enzyme Measures

One goal of this exploratory study is to describe enzyme levels overall, and
observe how they change (or don't), over time in our critically ill population.
We measure three enzymes on multiple days throughout the index hospitalization:

- AChE (acetylcholinesterase), units per liter (uL)
- AChE, units per gram of hemoglobin (ugHb)
- BChE (butyrylcholinesterase), units per liter

The figures below show both the overall distribution of *all* enzyme levels
during the entire hospitalization, and distributions by study day among patients
who are still alive and in the ICU at each time point.

```{r enzyme_overall, fig.height = 7, message = FALSE, warning = FALSE}
## We will use these arguments >=3 times, so time to make 'em a list
enzyme_args <- list(
  enzyme_var = c("cad_ache_ul", "cad_ache_ughb", "cad_bche_ul"),
  enzyme_title = c(
    "AChE, units/L", "AChE, units/gram hemoglobin", "BChE, units/L"
  )
)

## -- Calculate descriptive statistics for all enzymes, overall and by day -----
enzyme_desc_all <- daily_sub %>%
  dplyr::select(matches("cad\\_.*\\_u[lhgb]+$")) %>%
  skimr::skim_to_wide() %>%
  mutate(
    desc_string = glue(
      "N: {trimws(n)}\n",
      "Median (IQR): {trimws(median)} ({trimws(p25)}, {trimws(p75)})\n",
      "Mean (SD): {trimws(mean)} ({trimws(sd)})"
    )
  )

enzyme_desc_daily <- daily_sub %>%
  dplyr::select(cad_day, matches("cad\\_.*\\_u[lhgb]+$")) %>%
  group_by(cad_day) %>%
  skimr::skim_to_wide() %>%
  mutate(
    desc_string = glue(
      "N: {trimws(n)}\n",
      "Median (IQR): {trimws(median)} ({trimws(p25)}, {trimws(p75)})\n",
      "Mean (SD): {trimws(mean)} ({trimws(sd)})"
    )
  )

## -- Function to generate table of descriptives for a given enzyme ------------
enzyme_sumtable <- function(enzyme_var, enzyme_title){
  bind_rows(
    filter(enzyme_desc_all, variable == enzyme_var) %>%
      mutate(cad_day = 0),
    filter(enzyme_desc_daily, variable == enzyme_var)
  ) %>%
    mutate_if(is.character, trimws) %>%
    mutate_if(is.character, as.numeric) %>%
    mutate(
      cad_day_f = case_when(
        cad_day == 0 ~ "Overall",
        TRUE ~ paste("Day", cad_day)
      )
    ) %>%
    dplyr::select(cad_day_f, n, median, p25, p75, mean, sd)
}

## Create a single df with all descriptives
enzyme_descstats <- pmap(.l = enzyme_args, .f = enzyme_sumtable) %>%
  reduce(left_join, by = "cad_day_f")

## -- Data management for boxplots ---------------------------------------------
enzyme_boxdata <- bind_rows(
  daily_sub %>% mutate(cad_day = 0),
  daily_sub
) %>%
  ## Create factor versions of study day
  mutate(
    cad_day_f = factor(
      case_when(
        cad_day == 0  ~ 21,
        cad_day == 1  ~ 14,
        cad_day == 3  ~ 7,
        cad_day == 5  ~ 5,
        cad_day == 7  ~ 3,
        cad_day == 14 ~ 1,
        TRUE ~ as.numeric(NA)
      ),
      levels = c(1, 3, 5, 7, 14, 21),
      labels = c(paste("Day", c(14, 7, 5, 3, 1)), "Overall")
    )
  )

## -- Function to create boxplots ----------------------------------------------
enzyme_box <- function(enzyme_var, enzyme_title, df = enzyme_boxdata){
  bp <- ggplot(
    data = df, aes_string(x = "cad_day_f", y = enzyme_var, group = "cad_day_f")
  ) +
    geom_point(aes(colour = cad_day_f == "Overall"),
               alpha = 0.2, position = position_jitter(width = 0.1)) +
    geom_boxplot(
      aes(size = cad_day_f == "Overall"),
      fill = NA, outlier.shape = NA
    ) +
    scale_colour_manual(values = c("#B40F20", "#003D74"), guide = FALSE) +
    scale_size_manual(values = c(0.5, 1.0), guide = FALSE) +
    theme_bw() +
    coord_flip() +
    labs(
      y = enzyme_title,
      x = NULL
    ) +
    theme(
      axis.ticks.y = element_blank(),
      axis.text.y = element_text(face = "bold", size = 14)
    )
  
  return(bp)
}

## -- Function to create ridgeplots --------------------------------------------
enzyme_ridges <- function(enzyme_var, enzyme_title, df = enzyme_boxdata){
  rp <- ggplot(data = df, aes_string(y = "cad_day_f", x = enzyme_var)) +
    ggridges::geom_density_ridges(
      aes(fill = cad_day_f == "Overall"), alpha = 0.8
    ) +
    scale_fill_manual(values = c("#B40F20", "#003D74"), guide = FALSE) +
    theme_bw() +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    labs(x = enzyme_title)
  
  return(rp)
}

## -- Function to combine boxplots + ridgeplots for a single enzyme ------------
enzyme_combined <- function(bp, rp, cwidths = c(2, 1)){
  bp + rp + plot_layout(ncol = 2, widths = cwidths)
}

## -- Use purrr to make overall plots for all enzymes --------------------------
enzyme_boxplots <- map2(
  .x = c("cad_ache_ul", "cad_ache_ughb", "cad_bche_ul"),
  .y = c("AChE, units/L", "AChE, units/gram hemoglobin", "BChE, units/L"),
  .f = enzyme_box
)

enzyme_ridgeplots <- map2(
  .x = c("cad_ache_ul", "cad_ache_ughb", "cad_bche_ul"),
  .y = c("AChE, units/L", "AChE, units/gram hemoglobin", "BChE, units/L"),
  .f = enzyme_ridges
)

enzyme_plots <- map2(
  .x = enzyme_boxplots, .y = enzyme_ridgeplots, .f = enzyme_combined
)

## Old code; saving only till I get input from PIs on visual types
# ache_ul_overall <- ggplot() +
#   # ## Add shading in middle 80%
#   # geom_rect(
#   #   xmin = as.numeric(subset(enzyme_desc_all, variable == "cad_ache_ul")$p10),
#   #   xmax = as.numeric(subset(enzyme_desc_all, variable == "cad_ache_ul")$p90),
#   #   ymin = -Inf, ymax = Inf,
#   #   data = subset(enzyme_desc_all, variable == "cad_ache_ul"),
#   #   fill = "grey50", alpha = 0.2
#   # ) +
#   # ## Add shading in middle 50%
#   # geom_rect(
#   #   xmin = as.numeric(subset(enzyme_desc_all, variable == "cad_ache_ul")$p25),
#   #   xmax = as.numeric(subset(enzyme_desc_all, variable == "cad_ache_ul")$p75),
#   #   ymin = -Inf, ymax = Inf,
#   #   data = subset(enzyme_desc_all, variable == "cad_ache_ul"),
#   #   fill = "grey30", alpha = 0.2
#   # ) +
#   # geom_vline(
#   #   xintercept = as.numeric(subset(enzyme_desc_all, variable == "cad_ache_ul")$median),
#   #   linetype = "dotted"
#   # ) +
#   geom_histogram(
#     aes(x = cad_ache_ul), data = daily_sub, fill = "#B40F20"#, alpha = 0.6
#   ) +
#   geom_text(
#     aes(x = 8750, y = 40, label = desc_string),
#     hjust = 1,
#     size = 4,
#     data = enzyme_desc_all %>% filter(variable == "cad_ache_ul")
#   ) +
#   ## Can't get this to work well
#   # ggstance::geom_boxploth(aes(x = cad_ache_ul, y = 0), data = daily_sub, fill = NA) +
#   labs(
#     title = expression("AChE, "*mu*"L, All Study Days Combined"),
#     # subtitle = enzyme_desc_all %>%
#     #   filter(variable == "cad_ache_ul") %>%
#     #   pull(desc_string),
#     x = expression("AChE, "*mu*"L"),
#     y = "Frequency"
#   ) +
#     theme_minimal()
# 
# # ache_ul_overall
# 
# ache_ul_daily <- ggplot(data = daily_sub) +
#   facet_wrap(~ cad_day, ncol = 1) +
#   geom_histogram(
#     aes(x = cad_ache_ul), data = daily_sub, fill = "#B40F20"#, alpha = 0.6
#   ) +
#   geom_text(
#     aes(x = 8750, y = 10, label = desc_string),
#     hjust = 1,
#     size = 3,
#     data = enzyme_desc_daily %>% filter(variable == "cad_ache_ul")
#   ) +
#   labs(
#     title = expression("AChE, "*mu*"L, By Study Day"),
#     x = expression("AChE, "*mu*"L"),
#     y = NULL
#   ) +
#     theme_minimal()
# 
# # ache_ul_daily
# 
# # ache_ul_raw <- ggplot(data = daily_sub, aes(y = cad_ache_ul, x = dummy)) +
# #   geom_point(position = position_jitter()) +
# #   geom_boxplot(fill = NA) +
# #   coord_flip() +
# #   theme_minimal()
# # 
# # ache_ul_raw
# 
# ache_ul_overall + ache_ul_daily + plot_layout(nrow = 1, widths = c(2, 1))

```

## Visual{.tabset}

### AChE, units/L

```{r ache_ul_overall, fig.height = 7}
enzyme_plots[[1]]

```

### AChE, units/gram hemoglobin

```{r ache_ughb_overall, fig.height = 7}
enzyme_plots[[2]]

```

### BChE, units/L

```{r bche_ul_overall, fig.height = 7}
enzyme_plots[[3]]

```

## Tabular

```{r enzyme_desc}
kable(
  enzyme_descstats,
  format = "html",
  col.names = c(
    "Study Day",
    rep(c("N", "Median", "25th %ile", "75th %ile", "Mean", "SD"), 3)
  )
) %>%
  add_header_above(
    c(" " = 1,
      "AChE, units/L" = 6,
      "AChE, units/gram hemoglobin" = 6,
      "BChE, units/L" = 6)
  ) %>%
  mykablestyle(stripes = TRUE) %>%
  scroll_box(width = "700px")

```

# Enzymes vs Mental Status
