---
title: "CADUCEUS Preliminary Results"
subtitle: "Presentation at Kohler, April 2018"
author: "Jennifer Thompson, MPH; Supervisor: Rameela Chandrasekhar, PhD"
output:
  html_notebook:
    theme: flatly
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
---

CADUCEUS is an ancillary study enrolling patients from the MENDS2, MOSAIC, and
INSIGHT studies in progress at Vanderbilt University Medical Center, looking at
three enzyme levels (BChE, AChE, and AChe in hemoglobin) during the course of
critical illness. This preliminary analysis describes study results as of April
2018.

```{r setup, results = "hide"}
knitr::opts_chunk$set(message = FALSE)

suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(glue))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(captioner))
suppressPackageStartupMessages(library(rms))
suppressPackageStartupMessages(library(naniar))
suppressPackageStartupMessages(library(sparkline))

## Hmisc setup for nicely printing summaryM output
mu <- markupSpecs$html

## -- General helper functions (summary stats, formatting/printing) ------------
factor_tf <-
  function(x){ factor(as.numeric(x), levels = 1:0, labels = c("Yes", "No")) }
format_comma <- partial(format, big.mark = ",")
sum_na <- partial(sum, na.rm = TRUE)
q25 <- partial(quantile, probs = 0.25, na.rm = TRUE)
q50 <- partial(quantile, probs = 0.50, na.rm = TRUE)
q75 <- partial(quantile, probs = 0.75, na.rm = TRUE)
mean_na <- partial(mean, na.rm = TRUE)
sd_na <- partial(sd, na.rm = TRUE)
rndformat <- function(x, digits = 2){ format(round(x, digits), nsmall = digits) }
get_npct <- function(num, denom){
  sprintf("%s (%s%%)", num, round((num / denom) * 100))
}
describe_cont <- function(
  v_q50, v_q25, v_q75, v_mean, v_sd, dig_iqr = 0, dig_msd = 2
){
  sprintf(
    "%s [%s, %s]\\\n%s +/- %s",
    rndformat(v_q50, digits = dig_iqr),
    rndformat(v_q25, digits = dig_iqr),
    rndformat(v_q75, digits = dig_iqr),
    rndformat(v_mean, digits = dig_msd),
    rndformat(v_sd, digits = dig_msd)
  )
}

## Wrapper with my preferred options for printing summaryM objects
my_html <- function(sMobj, caption){
  html(
    sMobj,
    exclude1 = FALSE, long = TRUE, digits = 2, what = "%", npct = "both",
    prmsd = TRUE, brmsd = TRUE, middle.bold = TRUE,
    ## These options don't seem to be working.
    msdsize = mu$smaller2, outer.size = mu$smaller2, rowsep = TRUE,
    caption = caption
  )
}

## kable_styling wrapper to ensure all tables are consistently styled
mykablestyle <- function(obj, stripes = FALSE, ...){
  boptions <- c("hover", "responsive", "condensed", "bordered")
  if(stripes){ boptions <- c(boptions, "striped") }
  
  kable_styling(
    obj,
    bootstrap_options = boptions,
    full_width = FALSE,
    ...
  )
}

```

```{r captions}
## -- Table and figure captions ------------------------------------------------
table_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")

# table_nums(
#   name = "screening", caption = "Overview of Screening, Exclusions & Randomization"
# )
# 
# fig_nums(name = "exclusions", caption = "Study Exclusions")

```

```{r load_data}
## -- Source script that reads in, combines current data from all studies ------
source("combine_data.R")

```

# Description of Cohort

# Description of Enzyme Measures

# Enzymes vs Mental Status
